D3064R0: Bringing the Laws of Physics to Bear on the OOTA Problem

### 2024-01-11 ootavacuous.tex
Add section on multithreaded optimizations
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

@@

These tests might be two decades old, but they are still quite
relevant.

>>>>

\subsection{But What About Multi-Thread Optimizations?}
\label{sec:But What About Multi-Thread Optimizations?}

Although this paper mostly ignores multi-threaded implementations,
JMM Causality Test Cases 1, 8, 9, and 9a discuss inter-thread
analysis or constraints on thread scheduling.
This topic is therefore worth a short discussion.

In order to set context, it is important to note that an oracular C++
implementation could in theory read a program and its input, and then
perform a sequence of observable behaviors consistent with that program
running on that input without actually executing that program's code.
This admittedly extreme example shows that it is not unreasonable to
impose some real-world constraints on a real C++ implementation, and
limiting analysis to a single thread is one of the constraints that we
have chosen.

At the same time, it is worth calling out some optimizations based on
multi-threaded analysis that can be introduced without risk of
introducing OOTA cycles.

First, given a variable that the implementation can prove is always equal
to some constant, the implementation may act as if the program's uses of
that variable's value instead used that constant.

Second, given a pair of variables that the implementation can prove always
have a given algebraic relation to each other, the implementation can
act as if that program's uses of one of the variable's values instead
used the corresponding function of the other variable's value.
However, if the values of those variables are subject to change, there
must be some form of synchronization in effect to prevent accesses
to those variables' values during times when a given modification has
changed the value of one of the variables but not yet that of the other.
If there is no such synchronization, then this optimization cannot
be applied.

Third, given that the implementation can prove that one thread always
terminates before another thread starts, and if one or another of the
threads does nothing related to thread identity, the implementation
may act as if these two threads had been fused into a single thread.

Fourth and finally, if the implementation can prove that a given thread
never executes, the implementation may act as if code specific to that
thread had been omitted from the program.

Although these last two examples might appear to be a bit utopian,
the first two provide ample room for multi-threaded analysis.

@@

These tests might be two decades old, but they are still quite
relevant.\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}
